---
title: "P8130_Final_Reading"
author: "Yujing FU"
date: "2024-12-18"
output: html_document
---
```{r message=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(caret)
```

```{r}
# import dataset (whole data)
school <- read.csv("Project_1_data.csv", na = c("NA", ",", "")) |>
  janitor::clean_names() |>
  drop_na()

# dataset for math score
math <- school |>
  select(-reading_score, -writing_score)

# dataset for reading score
reading <- school |>
  select(-math_score, -writing_score)

# dataset for writing score
writing <- school |>
  select(-math_score, -reading_score)
```

```{r}
# whole school dataset transformed to indicator 0, 1, 2, 3, ...
school_transformed <- school |>
  mutate(across(everything(), as.factor)) |>
  mutate(across(where(is.factor), ~ as.integer(as.factor(.x)) - 1))

# whole school dataset transformed to indicator only containing 0 and 1
indicator_school <- as.data.frame(model.matrix(~ . - 1, data = school))

# math score indicator dataset
indicator_math <- indicator_school |>
  select(-reading_score, -writing_score)

# reading score indicator dataset
indicator_reading <- indicator_school |>
  select(-math_score, -writing_score)

# writing score indicator dataset
indicator_writing <- indicator_school |>
  select(-math_score, -reading_score)
```





```{r}
# model for reading_score
excluded_vars <- c("writing_score", "math_score")
predictors <- school[, !names(school) %in% excluded_vars]

# Define the response variable and the predictors
response <- school$reading_score
predictors <- predictors[, !names(predictors) %in% "reading_score"]

# Generate a formula with interaction terms for all pairwise combinations of predictors
predictor_names <- colnames(predictors)
interaction_terms <- paste(predictor_names, collapse = " + ")
formula <- as.formula(paste("reading_score ~ (", interaction_terms, ")^2"))

# Fit the linear regression model
model <- lm(formula, data = school)
summary(model)
```
From this overall linear regression model, we notice there are some significant covariates such as ??

```{r}
lm(reading_score ~ . + parent_educ*parent_marital_status+ lunch_type*practice_sport
, data = reading) |>
  summary()
```



```{r}
set.seed(1111)
model1 <- lm(reading_score ~ ., data = reading)

model2 <- lm(reading_score ~ . + parent_educ*parent_marital_status + lunch_type*practice_sport, data = reading)

anova_results <- anova(model1, model2)

print(anova_results)
```
Based on the significance condition of overall regression, we select two interaction term: `parent_educ*parent_marital_status` and `lunch_type*practice_sport`. Then we apply ANOVA to test whether adding the interaction term is better. The ANOVA P value = 0.04374 less than 0.05, indicating that the model with the addition of interaction terms significantly improves the fitting effect.

```{r}
# check the Multicollinearity
library(car)
vif(model2)
# all GVIF^(1/(2*Df)) are smaller than 5
```

```{r}
# check correlation problem
library(ggcorrplot)
ggcorrplot(cor_matrix, hc.order = TRUE, type = "lower")

```


```{r}
# 诊断图
par(mfrow = c(2, 2))
plot(model2)

# 检查残差正态性
shapiro.test(residuals(model2))
## not normal --> need to transform

# 检查多重共线性
library(car)
vif_values <- vif(model2)
print(vif_values)
```
```{r}
# Box-Cox
library(MASS)
boxcox_result <- boxcox(model2, lambda = seq(-2, 2, 0.1))

optimal_lambda <- boxcox_result$x[which.max(boxcox_result$y)]
cat("Optimal Lambda:", optimal_lambda, "\n")
```
```{r}
# Transform Y based on the optimal lambda = 1.434343
reading_trans <- reading |>
  mutate(reading_score_trans = (reading$reading_score^optimal_lambda - 1) / optimal_lambda)


# check normality
hist(reading_trans$reading_score_trans, main = "Transformed Reading Score", xlab = "Transformed Score")

```
```{r}
library(dplyr)
# transformed model
reading_trans =
  reading_trans |>  select(-reading_score)

model3 <- lm(school$reading_score_trans ~ . + parent_educ*parent_marital_status + lunch_type*practice_sport, data = reading)

summary(model3)

```



