---
title: "P8130_Final_Reading"
author: "Yujing FU"
date: "2024-12-18"
output: html_document
---
```{r message=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(caret)
```

```{r}
# import dataset (whole data)
school <- read.csv("Project_1_data.csv", na = c("NA", ",", "")) |>
  janitor::clean_names() |>
  drop_na()

# dataset for math score
math <- school |>
  select(-reading_score, -writing_score)

# dataset for reading score
reading <- school |>
  select(-math_score, -writing_score)

# dataset for writing score
writing <- school |>
  select(-math_score, -reading_score)
```

```{r}
# whole school dataset transformed to indicator 0, 1, 2, 3, ...
school_transformed <- school |>
  mutate(across(everything(), as.factor)) |>
  mutate(across(where(is.factor), ~ as.integer(as.factor(.x)) - 1))

# whole school dataset transformed to indicator only containing 0 and 1
indicator_school <- as.data.frame(model.matrix(~ . - 1, data = school))

# math score indicator dataset
indicator_math <- indicator_school |>
  select(-reading_score, -writing_score)

# reading score indicator dataset
indicator_reading <- indicator_school |>
  select(-math_score, -writing_score)

# writing score indicator dataset
indicator_writing <- indicator_school |>
  select(-math_score, -reading_score)
```





```{r}
# model for reading_score
excluded_vars <- c("writing_score", "math_score")
predictors <- school[, !names(school) %in% excluded_vars]

# Define the response variable and the predictors
response <- school$reading_score
predictors <- predictors[, !names(predictors) %in% "reading_score"]

# Generate a formula with interaction terms for all pairwise combinations of predictors
predictor_names <- colnames(predictors)
interaction_terms <- paste(predictor_names, collapse = " + ")
formula <- as.formula(paste("reading_score ~ (", interaction_terms, ")^2"))

# Fit the linear regression model
model <- lm(formula, data = school)
summary(model)
```
From this overall linear regression model, we notice there are some significant covariates such as ??

```{r}
lm(reading_score ~ . + parent_educ*parent_marital_status+ lunch_type*practice_sport
, data = reading) |>
  summary()
```



```{r}
model1 <- lm(reading_score ~ ., data = reading)

model2 <- lm(reading_score ~ . + parent_educ*parent_marital_status + lunch_type*practice_sport, data = reading)

anova_results <- anova(model1, model2)

print(anova_results)
```
Based on the significance condition of overall regression, we select two interaction term: `parent_educ*parent_marital_status` and `lunch_type*practice_sport`. Then we apply ANOVA to test whether adding the interaction term is better. The ANOVA P value = 0.04374 less than 0.05, indicating that the model with the addition of interaction terms significantly improves the fitting effect.

```{r}

# Load the library
library(corrplot)

excluded_vars <- c("math_score", "writing_score", "reading_score")
predictors <- school_transformed[, !names(school_transformed) %in% excluded_vars]
predictors_numeric <- predictors[, sapply(predictors, is.numeric)]

# Compute the correlation matrix
cor_matrix <- cor(predictors_numeric, use = "pairwise.complete.obs")  # Ensures missing values are handled

# Plot the correlation matrix using corrplot
corrplot(cor_matrix, 
         method = "color",        # Use color to display correlations
         type = "upper",          # Show only the upper triangle
         order = "hclust",        # Cluster variables with similar correlations
         addCoef.col = "black",   # Add correlation coefficients in black
         tl.col = "black",        # Text label color
         tl.srt = 45,             # Rotate axis labels
         col = colorRampPalette(c("blue", "white", "red"))(200)  # Color scale
)

```



```{r}
lm(math_score ~ . , data = math) |>
  summary()
```



